---
title: "Colour"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

```{r}
library(dplyr)
library(simplevis)
library(palmerpenguins)
```

### Overview

In simplevis, users adhere to the following rules for adjusting colour:

* Always define the colours to use via the `pal` argument 
* If colouring by a variable, use a `*_col()` or `*_col_facet()` function, and define the `col_var`
* If the `col_var` is numeric, the `col_method` of `bin` or `quantile`, and the `col_cuts` to use. 
* If `col_method` equals `bin`, then alternatively you can use the `col_pretty_n`argument to adjust number of pretty breaks. 

In simplevis, there is one colour concept, which generally includes all aspects of features that is relevant to the visualisation family type.

This is a consistent method across all functions, which is intended to simplify colouring.

### Always define the colours to use via the `pal` argument 

`simplevis` has default palettes that are based on the viridis and D3 colour palettes. 

```{r}
gg_point(penguins, 
         x_var = bill_length_mm, 
         y_var = body_mass_g)
```

But this default palette can be changed by providing a character vector of hex codes to the `pal` argument.

```{r}
gg_point(penguins, 
         x_var = bill_length_mm, 
         y_var = body_mass_g, 
         pal = "#da3490")

```

Users can get access to a large amount of colour palettes through the `pals` package.

### If colouring by a variable, use a `*_col()` or `*_col_facet()` function, and define the `col_var`

To colour by a variable, use a `*_col()` function and then define that variable to be coloured using the `col_var` argument.

Note the graph below uses a default pal for a non-numeric colour variable.

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species)
```

You can adjust the colours of the coloured variable using the pal argument as described earlier.

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             pal = c("#da3490", "#9089fa", "#47e26f"))
```

The `pals` package provides a large amount of colour palettes.

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             pal = pals::brewer.dark2(3))
```

### If colouring by a numeric variable, define the `col_method` and `col_cuts`

All `simplevis` `*_col()` and `*_col_facet()` functions support colouring by a categorical variable.

In addition, `sf`, `point`, `bar`, `hbar` and `tile` `*_col()` and `*_col_facet()` functions support colouring by a numeric variable.

You do this by specifying whether you want to do this by:

* defining whether the col_method is to be by `bin` or `quantile`
* defining a vector or col_cuts. These should be between 0 and infinity (`Inf`) for `bin` and between 0 and 1 for `quantile`

```{r}
plot_data <- ggplot2::diamonds %>% 
  slice_sample(prop = 0.01)

gg_point_col(plot_data, 
             x_var = carat, 
             y_var = price, 
             col_var = z,
             col_method = "bin")
```


```{r}
plot_data <- ggplot2::diamonds %>% 
  slice_sample(prop = 0.01)

gg_point_col(plot_data, 
             x_var = carat, 
             y_var = price, 
             col_var = z,
             col_method = "bin",
             col_cuts = c(0, 3, 4, Inf))
```

```{r}
plot_data <- ggplot2::diamonds %>% 
  slice_sample(prop = 0.01)

gg_point_col(plot_data, 
             x_var = carat, 
             y_var = price, 
             col_var = z,
             col_method = "quantile",
             col_cuts = c(0, 0.25, 0.5, 0.75, 1), 
             pal = pals::parula(6))
```

### If `col_method` equals `bin`, then alternatively you can use the `col_pretty_n`argument. 

The `col_pretty_n` argument uses the pretty algorithm to aim for roughly the specified number of pretty breaks.

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = flipper_length_mm, 
             col_method = "bin",
             col_pretty_n = 10)
```



