---
title: "Colour"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

```{r}
library(dplyr)
library(simplevis)
library(palmerpenguins)
```

### Overview

In simplevis, users define the following to adjust colours:

1. Always define the colours to use via the `pal` argument 
2. If colouring by a variable, use a `*_col` or `*_col_facet` function, and define the `col_var`
3. If the `col_var` is numeric, the `col_method` of `bin` or `quantile`, and the `col_cuts` to use.  

### 1. Always define the colours to use via the `pal` argument 

The colour palette can be changed from the default viridis colours by providing a character vector of hex codes to the `pal` argument.

```{r}
gg_point(penguins, bill_length_mm, body_mass_g, pal = "#e7298a")

```

Users can get access to a large amount of colour palettes through the `pals` package.

### 2. If colouring by a variable, use a `*_col` or `*_col_facet` function, and define the `col_var`

To colour by a variable, use a `*_col` function

```{r}
gg_point_col(penguins, bill_length_mm, body_mass_g, species, 
             pal = c("#1B9E77", "#D95F02", "#7570B3"))
```

### 3. If colouring by a numeric variable, define the `col_method` and `col_cuts`

All `simplevis` `*_col` and `*_col_facet` functions support colouring by a categorical variable.

In addition, `sf` and `point` `*_col` and `*_col_facet` functions support colouring by a numeric variable.

You do this by specifying whether you want to do this by:

* defining whether the col_method is to be by `bin` or `quantile`
* defining a vector or col_cuts. These should be between 0 and infinity (`Inf`) for `bin` and between 0 and 1 for `quantile`

```{r}
plot_data <- ggplot2::diamonds %>% 
  slice_sample(prop = 0.01)

plot_data

gg_point_col(plot_data, 
                 x_var = carat, 
                 y_var = price, 
                 col_var = z,
                 col_method = "quantile",
                 col_cuts = c(0, 0.25, 0.5, 0.75, 1))
```


```{r}
gg_point_col(plot_data, 
                 x_var = carat, 
                 y_var = price, 
                 col_var = z,
                 col_method = "bin",
                 col_cuts = c(0, 1, 2, 3, 4, 5, Inf))
```

