---
title: "Colour"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(simplevis)
```

### Overview

In simplevis, users define the following to adjust colours:

1. a `pal` argument of which colours to use 
2. a `col_var` argument of which variable is to be mapped to colours within an `_col` function
3. an `alpha` argument of what the opacity of the fill of colours within polygons is to be
4. If the `col_var` is numeric, the `col_method` of `bin` or `quantile`, and the `col_cuts` to use.  

### 2. Defining the `pal` to adjust which colours to use

The colour palette can be changed from the default viridis colours by providing a character vector of hex codes to the `pal` argument.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 6}
ggplot_point(iris, Sepal.Width, Sepal.Length, pal = "#e7298a")
```

Users can get access to a large amount of colour palettes through the `pals` package.

### 2. Defining a `col_var` to map a variable to colours within an `_col` function

To colour by a variable, use a `_col` function

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 6}
ggplot_point_col(iris, Sepal.Width, Sepal.Length, Species, 
                 pal = c("#1B9E77", "#D95F02", "#7570B3"))
```

### 3. Defining the `alpha` to adjust the opacity of the fill of polygons.

Emphasis can be given to the outline of polygons by using an `alpha` argument to add opacity to the fill of polygons.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 6}
plot_data <- ggplot2::diamonds %>%
  mutate(cut = stringr::str_to_sentence(cut)) %>%
  mutate(cut = factor(cut, levels = c("Fair", "Good", "Very good", "Premium", "Ideal"))) %>% 
  group_by(cut) %>%
  summarise(average_price = round(mean(price) / 1000, 1))

plot_data

ggplot_hbar(plot_data, average_price, cut, alpha = 0.25)
```

### 4. If colouring by a numeric variable, defining the `col_method` and `col_cuts`

All `simplevis` `_col` functions provide for colouring by a categorical variable.

In addition, `sf` and `point` functions provide for colouring by a numeric variable.

You do this by specifying whether you want to do this by:

* defining whether the col_method is to be by `bin` or `quantile`
* defining a vector or col_cuts. These should be between 0 and infinity (`Inf`) for `bin` and between 0 and 1 for `quantile`

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 6}
plot_data <- ggplot2::diamonds %>% 
  slice_sample(prop = 0.01)

plot_data

ggplot_point_col(plot_data, 
                 x_var = carat, 
                 y_var = price, 
                 col_var = z,
                 col_method = "quantile",
                 col_cuts = c(0, 0.25, 0.5, 0.75, 1))

ggplot_point_col(plot_data, 
                 x_var = carat, 
                 y_var = price, 
                 col_var = z,
                 col_method = "bin",
                 col_cuts = c(0, 1, 2, 3, 4, 5, Inf))
```

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 6}
ggplot_sf_col(example_sf_polygon, 
               col_var = density, 
               col_method = "bin", 
               col_cuts = c(0, 10, 50, 100, 150, 200, Inf), 
               col_labels_dp = 0)
```

### `simplevis` colouring vs `ggplot2` 

`simplevis` unites two `ggplot2` concepts into one to make things simpler for users:

* the colouring of points and lines (including outlines) 
* the colouring of the fill of polygons

These concepts are referred to as the `col` and the `fill` in ggplot2. So geom_point uses `col` and `geom_bar` uses fill (and col). 

To map colours to a variable:

* In `simplevis`: use a `_col` function and specify the `col_var`
* In `ggplot2`: add a `col` and/or `fill` argument within `aes` to the relevant `geom` layer

To select colours to use:

* In `simplevis`: use `pal` argument and specify a vector of colours
* In `ggplot2`: if no variable is <i>not</i> mapped to colours, then provide a vector of colours to fill and/or col arguments within the `geom` layer. Otherwise, if a variable is mapped to colours, then provide a vector of colours the values argument within `scale_fill_manual` and/or `scale_colour_manual`.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 6}
plot_data <- ggplot2::diamonds %>%
  mutate(cut = stringr::str_to_sentence(cut)) %>%
  mutate(cut = factor(cut, levels = c("Fair", "Good", "Very good", "Premium", "Ideal"))) %>% 
  group_by(cut) %>%
  summarise(average_price = round(mean(price) / 1000, 1))

plot_data

ggplot_hbar(plot_data, average_price, cut, 
            pal = viridis::viridis(4)[2], # in simplevis, pal is always a vector of colours
            title = "Average diamond price by cut and clarity", 
            x_title = "Average price ($US thousands)",  
            y_title = "Cut")

ggplot(plot_data) + 
  geom_bar(aes(x = average_price, y = fct_rev(cut)), 
           stat = "identity", 
           fill = viridis::viridis(4)[2], # in ggplot2, fill is a vector of colours when outside of aes
           width = 0.75) +
   scale_x_continuous(expand = c(0, 0), limits = c(0, 5)) +
   labs(title = "Average diamond price by cut", 
        x = "Average price ($US thousands)",  
        y = "Cut") +
  simplevis::theme_hbar() 
```


```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 6}
plot_data <- ggplot2::diamonds %>%
  mutate(cut = stringr::str_to_sentence(cut)) %>%
  mutate(cut = factor(cut, levels = c("Fair", "Good", "Very good", "Premium", "Ideal"))) %>% 
  group_by(cut, clarity) %>%
  summarise(average_price = round(mean(price) / 1000, 1))

plot_data

ggplot_hbar_col(plot_data, average_price, cut, col_var = clarity, # in simplevis, the col_var is the same as the fill/col within an aes in ggplot2
                pal = viridis::viridis(8), # in simplevis, pal is always a vector of colours
                position = "dodge",
                title = "Average diamond price by cut and clarity", 
                x_title = "Average price ($US thousands)", 
                y_title = "Cut")

ggplot(plot_data) + 
  geom_bar(aes(x = average_price, y = fct_rev(cut), fill = clarity), # in ggplot2, fill inside an aes is the same as the col_var in simplevis
           stat = "identity", 
           position = "dodge",
           width = 0.75) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 6)) +
  scale_fill_manual(values = viridis::viridis(8)) + # in ggplot2, the colours when fill is inside an aes are specified as a vector in scale_fill_manual
  guides(fill = guide_legend(rev = TRUE)) +
  labs(title = "Average diamond price by cut and clarity", 
       x = "Average price ($US thousands)",  
       y = "Cut",
       fill = NULL) +
  simplevis::theme_hbar() 
```

