---
title: "Ggplotly"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(simplevis)
```

### Overview

The `plotly::ggplotly` function works to convert ggplot objects to interactive html widgets.

The `simplevis` package provides:

* a `text_var` argument within ggplot wrapper functions for use as the tooltip
* a `mutate_text` function to create a tooltip easily based on all or a subset of variables in the data
* a `plotly_camera` function to turn off all widgets other than the camera
* `plotly_col_legend` to manipulate the order of plotly legends.

### Turning a ggplot into an html widget

The `plotly::ggplotly` function provides the ability to convert the ggplot to an interactive object. 

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7}
plot_data <- ggplot2::diamonds %>% 
  slice_sample(prop = 0.01)

plot_data

plot <- gg_point(plot_data, carat, price)

plotly::ggplotly(plot) 
```

### Creating nice tooltips

Use `mutate_text`, `text_var` and `plotly::ggplotly(tooltip = "text")` to make pretty tooltips.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7}
plot_data <- ggplot2::diamonds %>% 
  slice_sample(prop = 0.01) %>% 
  mutate_text()

plot_data %>% 
  select(text)

plot <- gg_point(plot_data, carat, price, text_var = text)

plotly::ggplotly(plot, tooltip = "text") 
```

You can select a subset of variables for your tooltip.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7}
plot_data <- ggplot2::diamonds %>% 
  slice_sample(prop = 0.01) %>% 
  mutate_text(c("color", "clarity", "depth"))

plot_data %>% 
  select(text)

plot <- gg_point(plot_data, carat, price, text_var = text)

plotly::ggplotly(plot, tooltip = "text")
```

### Turning off widgets other than the camera

The `plotly_camera` function turns off all widgets other than the camera to make a cleaner graph.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7}
plot_data <- storms %>%
  group_by(year) %>%
  summarise(average_wind = mean(wind))

plot <- gg_vbar(plot_data, year, average_wind)

plotly::ggplotly(plot) %>% 
  plotly_camera() 
```

### Controlling the colour legend.

Sometimes the ggplotly legend displays elements in a different order to how you want them. You can reverse them or specify a specific order using `plotly_col_legend`. Note, this only affects the legend elements.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7}
plot_data <- dplyr::sample_frac(ggplot2::diamonds, 0.05) 
 
plot <- gg_point_col(plot_data, carat, price, color)
 
plotly::ggplotly(plot)

plotly::ggplotly(plot) %>% 
   plotly_col_legend(rev = TRUE)
```

### `sf` maps

sf ggplots can also be converted to interactive ggplotly objects using the same method.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
plot_data <- example_sf_point %>% 
  mutate_text()

plot_data

plot <- gg_sf_col(plot_data, trend_category, text_var = text,
              borders = nz, 
              pal = c("#4575B4", "#D3D3D3", "#D73027"), 
              title = "Monitored trends, 2008-17")

plot

plotly::ggplotly(plot, tooltip = "text") %>% 
  plotly_camera()
```

```{r, message = FALSE, warning = FALSE, fig.height = 6, fig.width = 7}
col_cuts <- as.vector(quantile(example_sf_point$median, probs = c(0, 0.25, 0.5, 0.75, 1)))

plot_data <- example_sf_point %>% 
  mutate(quartile = case_when(median < col_cuts[2] ~ "Q1", 
                              median < col_cuts[3] ~ "Q2",
                              median < col_cuts[4] ~ "Q3",
                              TRUE ~ "Q4")) %>% 
  mutate_text()

plot <- gg_sf_col_facet(plot_data, trend_category, quartile, text_var = text,
                      borders = nz,
                      pal = c("#4575B4", "#D3D3D3", "#D73027"), 
                      title = "Trends by median quartile, 2017")

plot

plotly::ggplotly(plot, tooltip = "text") %>% 
  plotly_camera()
```

