---
title: "Ggplotly"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

```{r}
library(dplyr)
library(simplevis)
library(palmerpenguins)
```

### Overview

The `plotly::ggplotly` function works to convert ggplot objects to interactive html widgets.

The `simplevis` package provides:

* a `text_var` argument within `gg_*` wrapper functions for use as the tooltip
* a `mutate_text` function to create a tooltip easily based on all or a subset of variables in the data
* a `plotly_camera` function to turn off all widgets other than the camera
* `plotly_col_legend` to manipulate the order of plotly legends.

### Turning a ggplot into an html widget

The `plotly::ggplotly` function provides the ability to convert the ggplot to an interactive object. 

```{r}
plot <- gg_point_col(penguins, 
                     x_var = bill_length_mm, 
                     y_var = body_mass_g, 
                     col_var = species, 
                     font_family = "Helvetica")

plotly::ggplotly(plot) 
```

### Turning off widgets other than the camera

The `plotly_camera` function turns off all widgets other than the camera to make a cleaner graph.

```{r}
plot <- gg_point_col(penguins, 
                     x_var = bill_length_mm, 
                     y_var = body_mass_g, 
                     col_var = species, 
                     font_family = "Helvetica")

plotly::ggplotly(plot) %>% 
  plotly_camera()
  
```

### Creating nice tooltips

Use `mutate_text`, `text_var` and `plotly::ggplotly(tooltip = "text")` to make pretty tooltips.

A variable can be added to the `text_var` in the `gg_*` function. This variable is then used in the ggplotly tooltip when `tooltip = text` is added to the `ggplotly` function.

`simplevis` provides a `mutate_text` function which can produce a variable that is a string or variable names and values for a tooltip. Note this function converts column names to sentence case using the `snakecase::to_sentence_case` function. 

```{r}

plot_data <- penguins %>% 
  mutate_text()

plot <- gg_point_col(plot_data, 
                     x_var = bill_length_mm, 
                     y_var = body_mass_g, 
                     col_var = species, 
                     text_var = text, 
                     font_family = "Helvetica")

plotly::ggplotly(plot, tooltip = "text") %>% 
  plotly_camera()
```

The `mutate_text` function uses all variables in the dataset by defalut, but a subset can be used if desired. 

```{r}
plot_data <- penguins %>% 
  mutate_text(c("species", "flipper_length_mm"))

plot <- gg_point_col(plot_data, 
                     x_var = bill_length_mm, 
                     y_var = body_mass_g, 
                     col_var = species, 
                     text_var = text, 
                     font_family = "Helvetica")

plotly::ggplotly(plot, tooltip = "text") %>% 
  plotly_camera()
```

### Controlling the plotly colour legend order

Sometimes the ggplotly legend displays elements in a different order to how you want them. You can reverse them or specify a specific order using `plotly_col_legend`. Note, this only affects the legend elements.

```{r}
plot_data <- dplyr::sample_frac(ggplot2::diamonds, 0.05) 
 
plot <- gg_point_col(plot_data, 
                     x_var = carat, 
                     y_var = price, 
                     col_var = color, 
                     font_family = "Helvetica")
 
plotly::ggplotly(plot) %>% 
   plotly_camera() %>% 
   plotly_col_legend(order = c(2:7, 1))
```

### `sf` maps

sf ggplots can also be converted to interactive ggplotly objects using the same method.

```{r, fig.height = 4}
plot_data <- example_sf_point %>% 
  mutate_text()

plot <- gg_sf_col(plot_data, 
                  col_var = trend_category, 
                  text_var = text,
                  borders = nz, 
                  font_family = "Helvetica")

plotly::ggplotly(plot, tooltip = "text") %>% 
  plotly_camera()
```


```{r, fig.height = 6}
col_cuts <- as.vector(quantile(example_sf_point$median, probs = c(0, 0.25, 0.5, 0.75, 1)))

plot_data <- example_sf_point %>% 
  mutate(quartile = case_when(median < col_cuts[2] ~ "Q1", 
                              median < col_cuts[3] ~ "Q2",
                              median < col_cuts[4] ~ "Q3",
                              TRUE ~ "Q4")) %>% 
  mutate_text()

plot <- gg_sf_col_facet(plot_data, 
                        col_var = trend_category, 
                        facet_var = quartile, 
                        text_var = text,
                        borders = nz,
                        font_family = "Helvetica")

plotly::ggplotly(plot, tooltip = "text") %>% 
  plotly_camera()
```

