---
title: "Leaflet"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 4.5, 
  fig.width = 6
)
```

```{r}
library(simplevis)
```

### Overview

`leaflet` is a library that provides beautiful html widget maps. This allow use of basemaps, and therefore the ability to zoom in on features with the context that these provide. `simplevis` provides `leaflet_sf` and `leaflet_sf_col` wrapper functions to map sf objects. Arguments in these have been aligned to ggplot terminology to make things simple for the user.

As per the ggplot sf functions:

* Data must be an `sf` object 
* Data must be of `POINT`/`MULTIPOINT`, `LINESTRING`/`MULTILINESTRING`, or `POLYGON`/`MULTIPOLYGON` geometry type
* Data must have a coordinate reference system (CRS) defined
* No `x_var` and `y_var` variables are required

There is no facetting functionality provided for leaflet.

### Colouring 

Colouring is the same as that for graphs.

```{r}
leaflet_sf_col(example_sf_polygon, 
               col_var = density, 
               col_method = "quantile", 
               col_cuts = c(0, 0.25, 0.5, 0.75, 0.95, 1))
```


```{r}
leaflet_sf_col(example_sf_polygon, 
               col_var= density, 
               col_method = "bin", 
               col_cuts = c(0, 10, 50, 100, 150, 200, Inf), 
               title = "Modelled density, 2017")
```


### Popups and hover text

The clickable popup will default to a `leafpop::popupTable` of all variables, but popups can be adjusted to a subset of column using the `popup_vars_vctr` argument.

```{r}
leaflet_sf_col(example_sf_point, 
               col_var = trend_category, 
               popup_vars_vctr = c("site_id", "median"))
```

The hover label will default to the colour variable, but can be adjusted using the `label_var` variable. 

```{r}
leaflet_sf_col(example_sf_point, 
               col_var = trend_category, 
               label_var = site_id)
```

### Basemaps

Users have a `basemap` argument that defaults to "light", but there are other options.

```{r}
leaflet_sf(example_sf_point, 
           basemap = "dark")
```

### Labels

Labels are not as flexible as in the ggplot2 wrapper functions. You must provide a manual vector of breaks, and if you use a function it cannot access the colour variable breaks. Note you have only changed the colour legend labels, and so the hover labels and popups will remain based on the underlying data.

```{r}
leaflet_sf_col(example_sf_point, 
               col_var = trend_category, 
               col_labels = c("Better", "Same", "Worse"))
```

Sometimes a colouring method of "quantile" will provide too many digits in your legend labels. You can modify these by specifying a vector of these bins manually or using the `interval_labels` function.

```{r}
leaflet_sf_col(example_sf_polygon, 
               col_var = density, 
               col_method = "bin", 
               col_cuts = c(0, 10, 50, 100, 150, 200, Inf), 
               col_labels = interval_labels(c(0, 10, 50, 100, 150, 200, Inf)),
               title = "Modelled density, 2017")
```

### Adding additional layers

As a leaflet object is produced, you can add additional layers with leaflet - although this may effect popups and labels.

```{r}
leaflet_sf_col(example_sf_point, 
               col_var = trend_category) %>% 
  leaflet::addPolygons(data = nz, 
                       color = "#35B779", 
                       weight = 3, 
                       fillOpacity = 0, 
                       opacity = 1)
```

### leaflet basemap stack for shiny

A leaflet basemap stack is available for use in shiny apps. It defaults to the top layer being "light". You can set the bounds by adding a vector or bbox of bounds.

```{r}
leaflet_basemap(bounds = c(166.70047,-34.45676, 178.52966,-47.06345))
```

You can also specify the top layer.

```{r}
bb <- rnaturalearth::ne_countries(scale = "small", 
                                  country = "Indonesia", 
                                  returnclass = "sf") %>% 
   sf::st_bbox() 
 
leaflet_basemap(bounds = bb, top_layer = "satellite") 
```

