---
title: "Leaflet"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 4.5, 
  fig.width = 6
)
```

```{r}
library(simplevis)
library(stars)
set.seed(123456789)
```

### Overview

`leaflet` is a library that provides beautiful html widget maps. This allow use of basemaps, and therefore the ability to zoom in on features with the context that these provide. `simplevis` provides wrapper functions to map `sf` and ``stars` objects. Arguments in these have been aligned to ggplot terminology to make things simple for the user.

For the `sf` functions:

* Data must be an `sf` object 
* Data must be of `POINT`/`MULTIPOINT`, `LINESTRING`/`MULTILINESTRING`, or `POLYGON`/`MULTIPOLYGON` geometry type
* Data must have a coordinate reference system (CRS) defined
* No `x_var` and `y_var` variables are required

For the `stars` functions:

* Data must be a `stars` object 
* Data must have a coordinate reference system (CRS) defined
* No `x_var` and `y_var` variables are required

There is no facetting functionality provided for leaflet.

### Colouring 

Colouring is the same as that for graphs.

```{r}
leaf_sf_col(example_point, 
            col_var = trend_category)
```

```{r}
leaf_sf_col(example_polygon,  
            col_var = density)
```

```{r}
leaf_stars_col(example_stars, 
               col_var = nitrate, 
               col_method = "bin", 
               col_breaks_n = 4, 
               col_na_rm = TRUE)
```

```{r}
leaf_stars_col(example_stars, 
               col_var = nitrate, 
               col_method = "bin", 
               col_cuts = c(0, 500, 1000, 2000, Inf), 
               col_na_rm = TRUE)
```

```{r}
leaf_sf_col(example_polygon, 
            col_var = density, 
            col_method = "quantile", 
            col_breaks_n = 4)
```


```{r}
leaf_sf_col(example_polygon, 
            col_var = density, 
            col_method = "quantile", 
            col_cuts = c(0, 0.25, 0.5, 0.75, 0.95, 1))
```


```{r}
leaf_stars_col(example_stars, 
               col_var = nitrate, 
               col_method = "bin", 
               col_breaks_n = 7, 
               col_na_rm = TRUE)
```

### sf popups and hover text

The clickable popup will default to a `leafpop::popupTable` of all variables, but popups can be adjusted to a subset of column using the `popup_vars_vctr` argument.

```{r}
leaf_sf_col(example_point, 
            col_var = trend_category, 
            popup_vars_vctr = c("site_id", "median"))
```

The hover label will default to the colour variable, but can be adjusted using the `label_var` variable. 

```{r}
leaf_sf_col(example_point, 
            col_var = trend_category, 
            label_var = site_id)
```

### Basemaps

Users have a `basemap` argument that defaults to "light", but there are other options.

```{r}
leaf_sf(example_point, 
        basemap = "dark")
```

### Legend labels

Legend labels are not as flexible as in the ggplot2 wrapper functions. You can adjust by using the `col_labels` argument, and providing a manual vector of breaks. 

```{r}
leaf_sf_col(example_point, 
            col_var = trend_category, 
            col_labels = c("Better", "Same", "Worse"))
```

### Adding additional layers

As a leaflet object is produced, you can add additional layers with leaflet - although this may effect popups and labels.

```{r}
leaf_sf_col(example_point, 
            col_var = trend_category) %>% 
  leaflet::addPolygons(data = example_borders, 
                       color = "#35B779", 
                       weight = 3, 
                       fillOpacity = 0, 
                       opacity = 1)
```

### leaflet basemap stack for shiny

A leaflet basemap stack is available for use in shiny apps. It defaults to the top layer being "light". You can set the bounds by adding a vector or bbox of bounds.

```{r}
leaf_basemap(bounds = c(166.70047,-34.45676, 178.52966,-47.06345))
```

You can also specify the basemap top layer.

```{r}
bb <- rnaturalearth::ne_countries(scale = "small", 
                                  country = "Indonesia", 
                                  returnclass = "sf") %>% 
   sf::st_bbox() 
 
leaf_basemap(bounds = bb, basemap = "satellite") 
```

### More sophisticate modal popups in shiny

In shiny, you can create modal dialog popups in sf maps by:

* In `leaf_sf*()`, adding a layer id variable argument and turning the default popup to FALSE
* Adding `ObserveEvent` code to observe the click on a feature, and then show something in a modal dialog box.

Note that `input$map_marker_click$id` returns the value of the variable assigned to the `layer_id`. 

```{r, eval=F}
  leaf_draw <- function() {
    
    size_reactive <- ifelse(input$leaf_zoom < 6, 1.5,
             ifelse(input$leaf_zoom < 7, 2, 
                    ifelse(input$leaf_zoom < 8, 3, 4)))
    
    title <- "Monitored trends, 2008\u201317"
    
    leaf_sf_col(
      leaf_data(),
      col_var = trend_category,
      size_point = size_reactive,
      col_title = title, 
      layer_id_var = site_id,
      popup = FALSE
    )
  }
  
  observe({
    req(leaf_data())
    req(input$leaf_zoom) 
    
    withProgress(message = "Loading", {
      leaf_clear()
      leaf_draw()
    })
  })
  
  observeEvent(input$leaf_marker_click, {
    
    site_median <- leaf_data() %>% 
      filter(site_id == input$leaf_marker_click$id) %>% 
      pull(median)
    
    showModal(
      modalDialog(
        renderPlot(
          gg_histogram(example_point, median, 
                       title = paste0(input$leaf_marker_click$id, " compared to all sites"), 
                       theme = gg_theme(font = "Helvetica", size_title = 16,  size_body = 15, gridlines_h = T)) +
            ggplot2::geom_vline(ggplot2::aes(xintercept = site_median), col = "red")
        ),
        easyClose = TRUE
      )
    )
  })
  
```



