---
title: "Leaflet"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(simplevis)
```

### Overview

`leaflet` is a library that provides beautiful html widget maps. This allow use of basemaps, and therefore the ability to zoom in on features with the context that these provide. `simplevis` provides `leaflet_sf` and `leaflet_sf_col` wrapper functions to map sf objects. Arguments in these have been aligned to ggplot terminology to make things simple for the user.

As per the ggplot sf functions:

* Data must be an `sf` object 
* Data must be of `POINT`/`MULTIPOINT`, `LINESTRING`/`MULTILINESTRING`, or `POLYGON`/`MULTIPOLYGON` geometry type
* Data must have a coordinate reference system (CRS) defined
* No `x_var` and `y_var` variables are required

There is no facetting functionality provided for leaflet.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
leaflet_sf_col(data = example_sf_point, 
              col_var = median, 
              col_method = "quantile",
              col_cuts = seq(0, 1, 0.25),
              title = "Monitored medians, 2008-17")
```


```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
leaflet_sf_col(example_sf_polygon, density,
     col_method = "bin", 
     col_cuts = c(0, 10, 50, 100, 150, 200, Inf),
     col_labels_dp = 0,
     title = "Modelled density, 2017")
```

The popup will default to a `leafpop::popupTable` of all variables, but can be adjusted to a subset using the `popup_vars_vctr` argument.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
leaflet_sf_col(example_sf_point, trend_category, popup_vars_vctr = c("site_id", "median"))
```

The tooltip will default to the colour variable, but can be adjusted using the `text_var` variable.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
leaflet_sf_col(example_sf_point, trend_category, text_var = site_id)
```

### Basemaps

Users have a `basemap` argument that defaults to "light" but has other options.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
leaflet_sf(example_sf_point, basemap = "dark")
```

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
leaflet_sf(example_sf_point, basemap = "satellite")
```

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
leaflet_sf(example_sf_point, basemap = "street")
```

### shiny

A leaflet basemap is available for use in shiny apps. It defaults to the top layer being "light".

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
simplevis::leaflet_basemap() 
```

You can set the bounds by adding a vector or bbox of bounds.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
leaflet_basemap(bounds = c(166.70047,-34.45676, 178.52966,-47.06345))

bb <- rnaturalearth::ne_countries(scale = "small", country = "Indonesia", returnclass = "sf") %>% 
   sf::st_bbox() 
 
simplevis::leaflet_basemap(bounds = bb) 
```

### Adding additional layers

As a leaflet object is produced, you can add additional layers with leaflet.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
leaflet_sf_col(example_sf_point, 
               col_var = trend_category, 
               pal = c("#0D94A3", "#C4C4C7", "#AE4E51")) %>% 
  leaflet::addPolygons(data = nz, color = "#35B779", weight = 3, fillOpacity = 0, opacity = 1)
```
