---
title: "Scales"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(simplevis)
```

### Overview

`simplevis` has been built to take advantage of the auto-complete. 

Arguments that relate to the x scale start with `x_`, those that relate to the y scales start with `y_` and those that relate to the colour scale start with `col_`.

Likewise, arguments that relate to facetting start with `facet_`.

### Numeric scales

`simplevis` graphs default to including zero in numeric scales. 

Use `x_zero` and `y_zero` to change this.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
ggplot_point(iris, Sepal.Width, Sepal.Length,
               x_zero = FALSE, y_zero = FALSE)

```

Adjust the number of breaks for numeric x and/or y scales.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
ggplot_point(iris, Sepal.Width, Sepal.Length, 
               y_pretty_n = 10,
               x_pretty_n = 3)
```

Transform numeric x and y scales.

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
ggplot_point(iris, Sepal.Width, Sepal.Length,
               x_trans = "sqrt", 
               y_trans = "log10")
```

Balance a numeric scale so that it has equivalence between positive and negative values.

```{r}
plot_data <- storms %>%
  group_by(year) %>%
  summarise(average_wind = round(mean(wind), 2)) 

ggplot_vbar(plot_data, year, average_wind,
            y_balance = TRUE)
```

You can adjust x, y or colour scale labels using `x_labels`, `y_labels` or `col_labels` arguments, and functions from the `scales` package. 

```{r, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 7}
ggplot_point(iris, Sepal.Width, Sepal.Length,
               y_labels = scales::number_format(accuracy = 0.01, scale = 1/100),
               y_title = "Sepal length (m)")
```

Or you can use a vector of labels. Note the difference between a `x_title` and `the x_labels`

```{r}
ggplot_point(iris, Sepal.Width, Sepal.Length,
             x_labels = c("Zero", "One", "Two", "Three", "Four", "Five"), 
             x_title = "Sepal width (cm)")
```

### Discrete scales

`simplevis` automatically orders hbar graphs of character variables from biggest to smallest.

```{r, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 7}
plot_data <- ggplot2::diamonds %>%
  mutate(cut = as.character(cut)) %>% 
  group_by(cut) %>%
  summarise(average_price = mean(price)) 
  
ggplot_hbar(plot_data, average_price, cut)

```

Discrete scales can be reversed easily using the relevant `y_rev` or `x_rev` argument. 

```{r, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 7}
ggplot_hbar(plot_data, average_price, cut,
            y_rev = TRUE)
```

If there is an inherent order to the character variable that you want it to plot in, then you should convert the variable to a factor. 

```{r, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 7}
cut_levels <-  c("Fair", "Good", "Very Good", "Premium", "Ideal")

plot_data <- plot_data %>% 
 mutate(cut = factor(cut, levels = cut_levels)) 

ggplot_hbar(plot_data, average_price, cut)
```

### Colour scales

Add a colour title.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7}
plot_data <- ggplot2::diamonds %>%
  group_by(cut, clarity) %>%
  summarise(average_price = mean(price))

ggplot_hbar_col(plot_data, average_price, cut, clarity, col_title = "Clarity")

```

Reverse the palette.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7}
ggplot_hbar_col(plot_data, average_price, cut, clarity,
                pal_rev = TRUE)

```

Reverse the order of coloured bar pieces.

```{r, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7}
ggplot_hbar_col(plot_data, average_price, cut, clarity,
                col_rev = TRUE)

```

