---
title: "Shiny"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

```{r}
library(dplyr)
library(simplevis)
library(palmerpenguins)
```

### Overview

It can be challenging to build a `shiny` app with `ggplot` based graphs that works when viewed on mobile devices. `simplevis` offer methods to make ggplot based objects that work on either non-mobile or mobile devices.

### Working with quoted inputs

In shiny it is often easier to work with quoted variables. For ggplot wrapper function, wrap the variables in `!!sym`, as shown below. 

```{r}
gg_point_col(penguins, 
             x_var = !!sym("bill_length_mm"), 
             y_var = !!sym("body_mass_g"), 
             col_var = !!sym("species"))

```

### Mobile friendly apps

A difficulty with shiny is making shiny apps that display graphs that work on mobile phones in portrait mode.

They often look unreadable (in portrait), as:

* titles neeed to start from the far left of the plot, and be left-aligned
* titles need to be wrapped
* x scale needs the minimum of labels to not look clutterred
* legends need to be on the bottom
* font sizes need to be appropriate

`simplevis` provides mobile support for  plots through a `mobile` argument, which makes the above changes. 

```{r  fig.height = 5.5, fig.width = 2.5}
plot_data <- storms %>%
  group_by(year, status) %>%
  summarise(wind = mean(wind))

gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             mobile = T)

```

From the `simplevis` templates, the `mobileDetect` function in the `ui.R` and associated javascript in the `www/js/mobile.js` returns a TRUE or FALSE `input$isMobile` value if the user is on a mobile device. 

This `input$isMobile` TRUE or FALSE value can then be added to the `mobile` argument in the `simplevis` `gg_*` function in the `server.R`.

One reliable way to ensure the plot viewed by mobile users is in the `ui.R` to use conditional panels to provide a plotly object to non-mobile users and a plot object to mobile users.

This method is demonstrated in the `simplevis` templates.

### App templates

Users can access the `simplevis` functions by using the `shiny_template` functions for the applicable app, and then clicking on the `download_code` button to access a zip file of the code. 

```{r, eval=F}
shiny_template("template1") # a graph and table
shiny_template("template2") # a leaflet map, as well as graph and table
```

### Thanks!

Thanks to Gervasio Marchand for creating the code to identify whether users are on a mobile device as described [here](https://g3rv4.com/2017/08/shiny-detect-mobile-browsers)
