---
title: "Shiny"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(simplevis)
```

### Overview

It can be challenging to build a `shiny` app with `ggplot` based graphs that works when viewed on mobile devices. `simplevis` offer methods to make ggplot based objects that work on either non-mobile or mobile devices.

### Working with quoted inputs

In shiny it is often easier to work with quoted variables. For ggplot wrapper function, wrap the variables in `!!sym`, as shown below. 

```{r}
plot_data <- ggplot2::diamonds %>%
  sample_frac(0.05) %>%
  mutate(cut = stringr::str_to_sentence(cut))

gg_point_col_facet(plot_data, !!sym("carat"), !!sym("price"), !!sym("color"), !!sym("cut"))
```

### Mobile friendly apps

A difficulty with shiny is making shiny apps that display graphs that work on mobile phones in portrait mode.

They often look unreadable (in portrait), as:

* titles neeed to start from the far left of the plot, and be left-aligned
* titles need to be wrapped
* x scale needs the minimum of labels to not look clutterred
* legends need to be on the bottom
* font sizes need to be appropriate

`simplevis` provides mobile support for  plots through a `mobile` argument, which makes the above changes. 

```{r, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 3}
plot_data <- ggplot2::diamonds %>%
  mutate(cut = stringr::str_to_sentence(cut)) %>%
  group_by(cut) %>%
  summarise(average_price = mean(price)) 

gg_hbar(plot_data, average_price, cut,
            title = "Average diamond price by cut",
            x_title = "Average price ($US)",
            y_title = "Cut", 
            mobile = TRUE)
```

The default font size scales to look about right when viewed on a mobile device.

Build one app that:

* identifies a `input$isMobile` value of whether a user is on a mobile device or not through a mobileDetect function in the `ui.R` and some javascript in the `www/js/mobile.js`
* in the `server.R`, generate a `reactive` plot object incorporating an `mobile = input$isMobile` argument into `simplevis` ggplot function
* in the `server.R`, render a plot object and a plotly object
* in the ui.R, use conditional panels to provide the rendered plotly object to non-mobile users and the rendered plot object to mobile users

The reason why a plotly object is not provided to mobile users is that `ggplotly` does not implement all aspects of ggplot. This makes it hard to make it work nicely every time. A more reliable approach is to just provide a ggplot object instead to those users.

### App templates

This method is demonstrated in two template apps that are provided called `template1` and `template2`. 

Users can access these functions by using the `run_template` functions for the applicable app, and then clicking on the `download_code` button to access a zip file of the code. 

You can use these either to understand the method outlined above or to use as templates.

```{r, eval=F}
run_template("template1") # a graph and table
run_template("template2") # a leaflet map, as well as graph and table
```

If you use these as templates, the basic method to create an app is:

* run `run_template("template1")` or  `run_template("template2")` and download the code to use as a template
* In `get_data.R`, extract, process and save your data into the `data` subfolder, including a zip file for download
* In `make_app_vis.R`, draft your visualisations with dummy character inputs
* In `global.R`, read your data in, and create any vectors required
* In `ui.R`, add a app title
* In `ui.R`. add `radioButtons` and other widgets
* In `server.R`, add code within reactive plot_data and plot components, change any dummy character inputs to shiny user inputs. Add a `mobile = input$isMobile` specification to any simplevis graphs if you are looking to support mobile users as well as desktop
* In `server.R`, add code for map and table components, as applicable
* In `www/About.Rmd`, update as necessary
* If using google tag-manager, obtain a tag and replace `GTM-XXXXXXX` with it in the `www/js/tag-manager-js` file. 

### Thanks!

Thanks to Gervasio Marchand for creating the code to identify whether users are on a mobile device as described here:
https://g3rv4.com/2017/08/shiny-detect-mobile-browsers