---
title: "simplevis"
author: "David Hodge"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simplevis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

```{r}
library(simplevis)
```

*** Overview

As of simplevis 7.0.0, all ggplot2 wrapper functions have been removed. 

Users should not rely on these, and ideally use the  [{ggblanket}](https://github.com/davidhodge931/ggblanket) package instead (or alternatively rely on version 6.3.0 (`devtools::install_version("simplevis", "6.3.0")`)).

What remains in simplevis is the leaflet wrapper functions. These may or may not be supported in the future, so use at your own risk.

```{r}
leaf_sf_col(example_point, 
            col_var = trend_category)
```

Refer to the [{leaflet}](https://StatisticsNZ.github.io/simplevis/articles/simplevis.html) to get started.