---
title: "Simplevis"
author: "David Hodge"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simplevis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(simplevis)
library(palmerpenguins)
library(ggplot2)
```

### Purpose 

`simplevis` is a package of `ggplot2` and `leaflet` wrapper functions that aims to make visualisation easier with less brainpower required.

### Visualisation family types

`simplevis` supports the following families of visualisation type:

bar

```{r}
plot_data <- storms %>%
  group_by(year) %>%
  summarise(wind = mean(wind))

gg_bar(plot_data, 
       x_var = year, 
       y_var = wind)
```

point

```{r}
gg_point(iris, 
         x_var = Sepal.Width, 
         y_var = Sepal.Length)
```

line

```{r}
plot_data <- storms %>%
  group_by(year) %>%
  summarise(wind = mean(wind))

gg_line(plot_data, 
        x_var = year, 
        y_var = wind)
```

hbar (i.e horizontal bar)

```{r}
plot_data <- ggplot2::diamonds %>%
  group_by(cut) %>%
  summarise(price = mean(price))

gg_hbar(plot_data, 
        x_var = price, 
        y_var = cut)
```

tile

```{r}
plot_data <- penguins %>% 
  group_by(species, sex) %>% 
  summarise(bill_length_mm = round(mean(bill_length_mm, na.rm = TRUE), 1))

gg_tile_col(plot_data, 
            x_var = sex, 
            y_var = species, 
            col_var = bill_length_mm, 
            label_var = bill_length_mm) 
```

boxplot

```{r}
gg_boxplot(storms, 
           x_var = year, 
           y_var = wind)
```

density

```{r}
gg_density(penguins, 
           x_var = body_mass_g)
```

sf

```{r}
gg_sf_col(example_sf_point, 
          col_var = trend_category, 
          borders = nz)
```

stars

```{r}
library(stars)

gg_stars_col(example_stars,
             col_var = nitrate,
             col_na_rm = TRUE,
             borders = nz)
```

### Colouring, facetting, neither or both

Each visualisation family generally has four functions (except `tile`, which has two).

The function name specifies whether or not a visualisation is to be coloured by a variable (`*_col()`), facetted by a variable (`*_facet()`), or neither (`*()`) or both of these (`*_col_facet()`). 

Colouring by a variable means that different values of a selected variable are to have different colours. Facetting means that different values of a selected variable are to have their facet.

A `*()` function such `gg_point()` requires only a dataset, an x variable and a y variable. 

```{r}
gg_point(penguins, 
         x_var = bill_length_mm, 
         y_var = body_mass_g)
```

A `*_col()` function such `gg_point_col()` requires only a dataset, an x variable, a y variable, and a colour variable. 

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = sex)
```

A `*_facet()` function such `gg_point_facet()` requires only a dataset, an x variable, a y variable, and a facet variable. 

```{r}
gg_point_facet(penguins, 
               x_var = bill_length_mm, 
               y_var = body_mass_g, 
               facet_var = species)
```

A `*_col_facet()` function such `gg_point_col_facet()` requires only a dataset, an x variable, a y variable, a colour variable, and a facet variable. 

```{r}
gg_point_col_facet(penguins, 
                   x_var = bill_length_mm, 
                   y_var = body_mass_g, 
                   col_var = sex, 
                   facet_var = species)
```

Data is generally plotted with a stat of `identity`, which means data is plotted as is. 

For boxplot, there is adefault stat of `boxplot`, which means the `y_var` will be transformed to boxplot statistics. 

For density, the stat of the `x_var` based on the density prefixed arguments that inform the density calculation.

Generally, an `x_var` and a `y_var` is required. However, `y_var` is not required for `density*()` functions. Neither `x_var` or `y_var` are required for `gg_sf*()` (or `leaflet_sf*()`) functions.

### Titles

Defaults titles are: 

 * no title, subtitle or caption 
 * x, y and colour titles are converted to sentence case using the `snakecase::to_sentence_case` function. 

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species)
```

You can customise titles with `title`, `subtitle`, `x_title`, `y_title` and `caption` arguments. 

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "Adult penguin mass by bill length and species",
             subtitle = "Palmer station, Antarctica",
             x_title = "Bill length (mm)", 
             y_title = "Body mass (g)",
             col_title = "Penguin species")
```

You can also request no x_title using `x_title = ""` or likewise for `y_title` and `col_title`.

### Colour palettes

Change the colour palette by supplying a vector of colours to the `pal` argument. 

```{r}
gg_point(iris, 
         x_var = Sepal.Width, 
         y_var = Sepal.Length, 
         pal = "#e7298a")
```

### Scales

simplevis makes it easy to make easy scale transformations.

These use consistent prefixes based on `x_*`, `y_*`, `col_*` or `facet_*`, and as such the autocomplete can help identify what you need.

Some examples of transformations available are:

* `*_na_rm` to quickly not include NA observations
* `*_labels` to adjust labels for any x, y, col or facet scale
* `*_zero` to start at zero for numeric x or y scales 
* `*_pretty_n` for the number of numeric bins of breaks for the x or y scale to aim for 
* `*_rev` to reverse the order of categorical x, y or col scales in bars 
* `*_gridlines_minor` whether to add minor gridlines.
* `*_expand` to add padding to an x or y scale.

```{r}
plot_data <- storms %>%
  group_by(year, status) %>%
  summarise(wind = mean(wind))

gg_bar_col(plot_data, 
        x_var = year, 
        y_var = wind, 
        col_var = status,
        position = "stack",
        x_pretty_n = 4,
        x_labels = function(x) stringr::str_sub(x, 3, 4),
        y_labels = function(x) scales::comma(x, accuracy = 0.1), 
        y_zero = T, 
        y_pretty_n = 10,
        y_gridlines_minor = T,
        y_expand = ggplot2::expansion(mult = c(0.025, 0.025)))
```

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = sex, 
             col_na_rm = TRUE)
```

`bar*()` and `hbar*()` plots support a `"stack"` position as well as the default `"dodge"`.

```{r}
plot_data <- penguins %>% 
  group_by(sex, species) %>% 
  summarise(count = n())

gg_hbar_col(plot_data, 
        x_var = count, 
        y_var = species, 
        col_var = sex,
        position = "stack")
```

### `sf` and `stars` maps

`simplevis` provides `sf` and `stars`  maps.

`sf` maps are maps of point, line or polygon features.

`stars` maps are maps of arrays (i.e. grids).

`sf` functions work in the same way as the ggplot2 graph functions, but with the following  differences:

* Data must be an `sf` object 
* Data must be of `POINT`/`MULTIPOINT`, `LINESTRING`/`MULTILINESTRING`, or `POLYGON`/`MULTIPOLYGON` geometry type
* Data must have a coordinate reference system (CRS) defined
* No `x_var` and `y_var` variables are required
* Borders can added to maps by providing an `sf` object to the `borders` argument. 

`stars` functions work in the same way as the ggplot2 graph functions, but with the following  differences:

* Data must be a `stars` object 
* Data must have a coordinate reference system (CRS) defined
* No `x_var` and `y_var` variables are required
* Borders can added to maps by providing an `sf` object to the `borders` argument. 

The following example objects are provided withing the package for learning purposes: `example_sf_point`, `example_sf_polygon` and `example_stars`.

The borders argument allows for the user to provide an sf object as context to the map (e.g. a coastline or administrative boundaries). An sf object of the New Zealand coastline has been provided to illustrate how this works.

```{r}
gg_sf_col(example_sf_point, 
          col_var = median, 
          col_method = "quantile",
          col_cuts = c(0, 0.25, 0.5, 0.75, 1),
          borders = nz)
```

```{r}
gg_stars_col(example_stars,
             col_var = nitrate,
             col_method = "quantile",
             col_cuts = c(0, 0.05, 0.25, 0.5, 0.75, 0.95, 1),
             col_na_rm = TRUE,
             borders = nz)
```

### leaflet wrappers

simplevis also provides a `leaflet` wrapper functions for `sf` and `stars` objects. 

```{r}
leaflet_sf_col(example_sf_point, 
               col_var = trend_category)
```


```{r}
library(stars)

leaflet_stars_col(example_stars, 
                  col_var = nitrate, 
                  col_method = "quantile", 
                  col_cuts = c(0, 0.05, 0.25, 0.5, 0.75, 0.95, 1),
                  col_na_rm = TRUE)

```

### Supported variable classes

variable types supported by the different groups of functions are outlined below.

A stat of `identity` refers to the value being plotted as it is. A stat of `boxplot` refers to boxplot statistics being calculated from the data, and these plotted.

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7}
tibble::tribble(
  ~family, ~data, ~x_var, ~y_var, ~col_var, ~facet_var, ~stat,
  "bar", "tibble or data.frame", "Any*", "Numeric", "Categorical or numeric", "Categorical", "identity",
  "hbar", "tibble or data.frame", "Numeric", "Any*", "Categorical or numeric", "Categorical", "identity",
  "line", "tibble or data.frame", "Any", "Numeric", "Categorical", "Categorical", "identity",
  "point", "tibble or data.frame", "Any", "Numeric", "Categorical or numeric", "Categorical", "identity",
  "density", "tibble or data.frame", "Numeric", NA, "Categorical",  "Categorical", "density",
  "boxplot", "tibble or data.frame", "Any*", "Numeric", "Categorical", "Categorical", "boxplot (or identity)",
  "tile",  "tibble or data.frame", "Categorical", "Categorical", "Categorical or numeric", "Categorical", "identity",
  "sf", "sf", NA, NA, "Categorical or numeric", "Categorical", "identity",
  "stars", "stars", NA, NA, "Categorical or numeric", NA, "identity",
  ) %>% 
  DT::datatable()
```

Where:

* Categorical refers to character, factor, or logical classes.
* Numeric refers to double or integer classes.
* Any* refers that if a numeric, date or datetime variable, values must be bins that are mutually exclusive and equidistant. 

### Working with the pipe

`simplevis` functions work with the pipe.

```{r}
penguins %>% 
  gg_density_col(x_var = body_mass_g, 
                 col_var = species)
```

### Output objects and adding layers

All `gg_*` and `leaflet_*` wrapper functions produce ggplot or leaflet objects.

This means layers can be added to the functions in the same way you would a ggplot2 or leaflet object.

Note you need to add _all_ aesthetics to any additional layers.

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species) +
  geom_smooth(aes(x = bill_length_mm, y = body_mass_g, col = species))

```

This means you can facet by more than one variable, provided that you are not using a position of "stack". 

```{r, fig.height = 4.5}
plot_data <- penguins %>% 
  group_by(species, sex, island) %>% 
  summarise(body_mass_g = mean(body_mass_g, na.rm = TRUE))

gg_bar(plot_data, 
       x_var = sex, 
       y_var = body_mass_g, 
       width = 0.66, 
       x_na_rm = TRUE, 
       y_pretty_n = 3) +
  facet_grid(rows = vars(species), 
             cols = vars(island), 
             labeller = as_labeller(snakecase::to_sentence_case))

```

### ggplotly interactive visualisation

All ggplot objects can be converted into interactive html objects using ggplotly. You can simply wrap the plot object in `plotly::ggplotly()`. 

The `plotly_camera` function removes plotly widgets other than the camera to keep things tidy.

```{r, eval = FALSE}
plot <- gg_point_col(penguins, 
                     x_var = bill_length_mm, 
                     y_var = body_mass_g, 
                     col_var = species) 

plotly::ggplotly(plot) %>% 
  plotly_camera()
```

`simplevis` also offers more customisability for making tooltips(i.e. hover values) in ggplotly (i.e. hover values). 

A variable can be added to the `text_var` in the `gg_*` function. This variable is then used in the ggplotly tooltip when `tooltip = text` is added to the `ggplotly` function.

`simplevis` provides a `mutate_text` function which can produce a variable that is a string or variable names and values for a tooltip. Note this function converts column names to sentence case using the `snakecase::to_sentence_case` function. 

The `mutate_text` function uses all variables in the dataset by defalut, but a subset can be used if desired. 

```{r, eval = FALSE}
plot <- gg_point_col(penguins, 
                     x_var = bill_length_mm, 
                     y_var = body_mass_g, 
                     col_var = species) 

plotly::ggplotly(plot) %>% 
  plotly_camera()

plot_data <- penguins %>% 
  mutate_text()

plot <- gg_point_col(plot_data, 
                     x_var = bill_length_mm, 
                     y_var = body_mass_g, 
                     col_var = species, 
                     text_var = text, 
                     font_family = "Helvetica")

plotly::ggplotly(plot, tooltip = "text") %>% 
  plotly_camera()
```

### Further information

For further information, see the articles on the [website](https://statisticsnz.github.io/simplevis/).