---
title: "Themes"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

```{r setup}
library(simplevis)
library(tidyverse)
library(palmerpenguins)
```

From version 5.1.0, `simplevis` will support all ggplot2 themes with all `gg_*` plot functions providing a theme argument. 

All graphs will have the same default undelying `simplevis` themes.

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle")
```

A `ggtheme` function will be provided to quickly create a theme. 

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle", 
             theme = gg_theme(gridlines = "horizontal", 
                              title_size = 20)) 
```

This has some flexibility.

```{r}
custom_theme <- gg_theme(
  family = "lato",
  title_pal = "darkgreen",
  title_face = "plain",
  title_size = 15,
  subtitle_pal = "lightgreen",
  subtitle_face = "bold",
  subtitle_size = 16,
  body_pal = "darkblue",
  body_face = "italic",
  body_size = 9,
  gridlines = "vertical",
  gridlines_pal = "pink", 
  gridlines_size = 0.7,
  ticks_pal = "red",
  ticks_size = 1)

gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle", 
             theme = custom_theme) 
```

Users can also create their own custom themes using the `ggplot2::theme` function, or use other themes that other packages provide.

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle",
             theme = ggplot2::theme_grey())
```

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle",
             col_title = "",
             theme = hrbrthemes::theme_ipsum())
```

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle",
             col_title = "",
             pal = ggthemes::economist_pal(fill = F)(3),
             theme = ggthemes::theme_economist())
```
