---
title: "Themes"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

```{r setup}
library(simplevis)
library(tidyverse)
library(palmerpenguins)
```

From version 5.1.0, `simplevis` will support all ggplot2 themes with all `gg_*` plot functions providing a theme argument. 

All graphs will have the same default undelying `simplevis` themes.

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle")
```

A `ggtheme` function will be provided to quickly create a theme. 

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle", 
             theme = gg_theme(gridlines = "horizontal", 
                              size_title = 20)) 
```

This has some flexibility.

```{r}
custom_theme <- gg_theme(
  family = "lato",
  size_title = 15,
  size_subtitle = 16,
  size_body = 9,
  size_axis = 1,
  size_gridlines = 0.7,
  size_ticks = 1,
  face_title = "plain",
  face_subtitle = "bold",
  face_body = "italic",
  pal_title = "darkgreen",
  pal_subtitle = "lightgreen",
  pal_body = "darkblue",
  pal_axis = "grey",
  pal_ticks = "red",
  pal_gridlines = "pink", 
  gridlines = "vertical")

gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle", 
             theme = custom_theme) 
```

Users can also create their own custom themes using the `ggplot2::theme` function, or use other themes that other packages provide.

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle",
             theme = ggplot2::theme_grey())
```

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle",
             col_title = "",
             theme = hrbrthemes::theme_ipsum())
```

```{r}
gg_point_col(penguins, 
             x_var = bill_length_mm, 
             y_var = body_mass_g, 
             col_var = species, 
             title = "A nice long title", 
             subtitle = "And a subtitle",
             col_title = "",
             pal = ggthemes::economist_pal(fill = F)(3),
             theme = ggthemes::theme_economist())
```
