---
title: "Titles"
author: "David Hodge"
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3.5, 
  fig.width = 6
)
```

```{r}
library(dplyr)
library(simplevis)
library(palmerpenguins)
set.seed(123456789)
```

### Default titles

Default titles in simplevis try to provide a polished quick graph with minimal code by:

* having no title, subtitle or caption 
* converting x, y and colour titles to sentence case using the `snakecase::to_sentence_case` function.

```{r}
gg_point_col(penguins, bill_length_mm, body_mass_g, species)
```

### Adjusting titles

Depending on column names and the purpose of the visualisation, the sentence case transformation of column names may not be appropriate titles, and you will need to change them manually.

You can do this using the `title`, `subtitle`, `x_title`, `y_title`, `col_title` and `caption` arguments. 

```{r}
gg_point_col(penguins, bill_length_mm, body_mass_g, species, 
             title = "Adult penguin mass by bill length and species",
             subtitle = "Palmer station, Antarctica",
             x_title = "Bill length (mm)", 
             y_title = "Body mass (g)",
             col_title = "Penguin species")
```

###  Turning off x, y and colour titles

If you want no x, y or colour title, you need to use `x_title = ""`, `y_title = ""` or `col_title = ""` as applicable.

```{r}
gg_point_col(penguins, bill_length_mm, body_mass_g, species, 
             x_title = "")
```




